---
title: "Statistical Rethinking Ch. 3 Exercises"
author: Cassandra
output: 
  html_document: 
    keep_md: yes
---

```{r message = FALSE}
library(rethinking)
```

# Easy

```{r}
p_grid <- seq(from = 0, to = 1, length.out = 1000)
prior <- rep(1, 1000)
likelihood <- dbinom(6, size = 9, prob = p_grid)
posterior <- likelihood * prior
set.seed(100)
samples <- sample(p_grid, prob = posterior, size = 1e4, replace = TRUE)

# plot posterior distribution
plot(p_grid, posterior, type = "b")
```

## 3E1 - How much posterior probability lies below *p* = 0.2?  

```{r}
sum(samples < 0.2) / 1e4
```
0.05% of posterior probability is below *p* = 0.2.

## 3E2 - How much posterior probability lies above *p* = 0.8?

```{r}
sum(samples > 0.8) / 1e4
```
11.17% of posterior probability is above *p* = 0.8.

## 3E3 - How much posterior probability lies between *p* = 0.2 and *p* = 0.8?

```{r}
sum(samples > 0.2 & samples < 0.8) / 1e4
```
88.78% of posterior probability is between *p* = 0.2 and *p* = 0.8.

## 3E4 - 20% of the posterior probability lies below which value of *p*?

```{r}
quantile(samples, 0.2)
```
20% of the posterior probability is below *p* = 0.52.

## 3E5 - 20% of the posterior probability lies above which value of *p*?

```{r}
quantile(samples, 0.8)
```
20% of the posterior probability is above *p* = 0.76.

## 3E6 - Which values of *p* contain the narrowest interval equal to 66% of the posterior probability?

```{r}
HPDI(samples, prob = 0.66)
```
The narrowest interval equal to 66% of the posterior probability is from *p* = 0.52 to *p* = 0.78.

## 3E7 - Which values of *p* contain 66% of the posterior probability, assuming equal posterio probability both below and above the interval?

```{r}
PI(samples, prob = 0.66)
```
Assuming equal posterior probability below and above the interval, 0.5 < *p* < 0.77 contains 66% of the posterior probability.

# Medium

## 3M1 - Suppose the globe tossing data turned out to be 8 water in 15 tosses. Construct the posterior distribution, using grid approximation. Use the same flat prior as before.

```{r}
p_grid <- seq(from = 0, to = 1, length.out = 1000)
prior <- rep(1, 1000)
likelihood <- dbinom(8, size = 15, prob = p_grid)
posterior <- likelihood * prior
posterior <- posterior / sum(posterior)

# graph posterior distribution
plot(p_grid, posterior, type = "b")
```


## 3M2 - Draw 10,000 samples from the grid approximation from above. Then use the samples to calculate the 90% HPDI for *p*.

```{r}
# draw 10,000 samples
samples <- sample(p_grid, prob = posterior, size = 1e4, replace = TRUE)

# calculate 90% HPDI
HPDI(samples, prob = 0.9)
```
The 90% HPDI for *p* is 0.338 < *p* < 0.732.

## 3M3 - Construct a posterior predictive check for this model and data. This means simulate the distribution of samples, averaging over the posterior uncertainty in *p*. What is the probability of observing 8 water in 15 tosses?

```{r}
w <- rbinom(1e4, size = 15, prob = samples)
simplehist(w)

# probability of observing 8 water in 15 tosses
sum(w == 8) / 1e4
```
There is a 14% probability of observing 8 water in 15 tosses.

## 3M4 - Using the posterior distribution constructed from the new (8/15) data, now calculate the probability of observing 6 water in 9 tosses.

```{r}
w <- rbinom(1e4, size = 9, prob = samples)
simplehist(w)

# probability of observing 6 water in 9 tosses
sum(w == 6) / 1e4
```
There is a 17% probability of observing 6 water in 9 tosses.
